\racets: Policy{-}Agnostic Programming in Racket

### 2018-07-23 paper/paper.tex
pass on section 2

In this paper we present \racets, an implementation of policy-agnostic
programming in Racket via macros. \racets provides facilities for
creating policies and faceting secure data with those policies. \racets
also extends several core forms in Racket to work with faceted
values (our implementation is detailed in Section~\ref{sec:implementation}). We
have measured the overhead of \racets in several microbenchmarks, and
have used \racets to implement a small server-based
board-game (detailed in Section~\ref{sec:overview}). We see \racets as a
promising prototype for policy-agnostic programming in Racket, and
conclude with discussion of future directions in Section~\ref{sec:conclusion}.

@@

To introduce our setting we present the implementation of Battleship,
a small guessing-based board game, in \racets (this section presents a
distilled version of our case study in
Section~\ref{sec:case-study}). In this game each player has a
grid-based board on which they place tiles (or ``ships''). The players
hide their boards from each other as play progresses in rounds. Each
turn a player guesses the position of a tile on the other player's
board. If the guess is successful that tile is removed from the
board. Play ends once one player's board has no remaining tiles, at
which point that player loses.

We implement game boards as lists of cons cells representing the
\blue{$(x,y)$} coordinates of ships. Board creation is simply the empty
list, and adding a piece is done via \lstinline|cons|:

@@

Although \code{mark-hit} will operate on sensitive data (the game
boards), it is written without any special machinery to maintain the
secrecy of \code{board}. Protecting data {w.r.t.} policies is instead handled
automatically and implicitly by a runtime monitor.
When Alice and Bob want to play a game, they both create a
\emph{label} to protect their data. A label is a predicate that takes
an argument and, if it returns true, reveals a secret. Alice's label
is used to annotate whatever data she wants to be kept
secret. Supposing Alice chooses to be player 1, she will use the
following label:

@@

Bob would use a similar label (but with 2 instead of 1). At runtime,
the \code{label} form creates a label \blue{$\ell_{A}$} and returns it
to the binding for \code{alice-label}. When Alice wants to protect a
value, she creates a facet annotated with her label and two
\emph{branches}. The positive (left) branch represents the value as it
should appear to her, and the negative (right) to everyone else:

@@

In the above example, Alice uses \blue{$\star$} (lazy failure) to
represent that others should see nothing if they observe the data. In
other applications, she might choose a more sensible default value to
reveal to others. She may even want to create a \emph{nested}
facet. For example, in a social-networking application she may want a
nested facet consisting of two labels for
\blue{$\ell_{\text{Friends}}$} and \blue{$\ell_{\text{Family}}$}. She
would present three views of her social-media profile: \bmth{p_1} to
her family, containing her phone number and other contact information,
\bmth{p_2} to her friends showing her interests, and \bmth{p_3} to
everyone else, showing only her name and email.

@@

As play of the game progresses, Alice and Bob both make guesses, and a
driver calls the function \code{mark-hit} with each of their (faceted)
game boards. However, because Alice and Bob's game boards are both
facets, \code{mark-hit} cannot be immediately applied, as the argument
\code{board} is a facet. Faceted execution ``splits'' the execution of
the function on faceted arguments, running it first on the positive
branch, then again on the negative branch. Finally, the results of
each branch are merged to produce a facet.

@@

Because the applied function can be stateful, faceted execution also
records the current privilege level in a \emph{program counter}. The
program counter is used to build facets when writes are made to the
store inside of a privileged context. We expand upon these subtleties
in Section~\ref{sec:semantics}, where we present a full semantics for
faceted execution.

\kris{Stuff about obs here..}

@@

\begin{figure}
\begin{displaymath}
  \begin{array}{lcrcl}
    \blue{c} & \blue{\in} & \blue{const} & \blue{::=} & \blue{() \mid \code{true} \mid \code{false} \mid \ldots} \\
    \blue{x} & \blue{\in} & \blue{var} & \blue{::=} & \blue{\ldots}  \\
    \blue{e} & \blue{\in} & \blue{exp} & \blue{::=} & \blue{c} \mid \blue{x} \\
    & & & \blue{\mid} & \blue{(\lambda ~ (x) ~ e) \mid (e ~ e)} \\
    & & & \blue{\mid} & \blue{(\code{box} ~ e) \mid (\code{unbox} ~ e) \mid (\code{set!} ~ e ~ e)} \\
    & & & \blue{\mid} & \blue{(\code{let-label} ~ \ell ~ e ~ e)} \\
    & & & \blue{\mid} & \blue{(\code{facet} ~ e ~ e ~ e)} \\
    & & & \blue{\mid} & \blue{(\code{obs} ~ e ~ e ~ e)} \\
  \end{array}
\end{displaymath}
\label{fig:syntax}
\caption{Syntax of \lang}
\end{figure}

>>>>

In this paper we present \racets, an implementation of policy-agnostic
programming in Racket via macros. \racets provides facilities for
creating policies and faceting secure data with those policies. \racets
also extends several core forms in Racket to work with faceted
values (our implementation is detailed in Section~\ref{sec:implementation}). We
have measured the overhead of \racets in several microbenchmarks, and
have used \racets to implement a small server-based
board-game (detailed in Sections~\ref{sec:overview} and~\ref{sec:implementation}). We see \racets as a
promising prototype for policy-agnostic programming in Racket, and
conclude with discussion of future directions in Section~\ref{sec:conclusion}.

@@

To introduce faceted execution more concretely, we present the implementation
of Battleship, a small guessing game, in \racets (this section presents a
distilled version of our case study in
Section~\ref{sec:implementation}). In this game each player has a
private board of coordinates, at which they place ``ships''. The players
hide their boards from each other as play progresses in rounds. Each
turn a player guesses the position of a ship on the other player's
board. If the guess is successful the tile is removed from the
board and a hit is declared publically. Play ends once one player's board
has no remaining tiles, at which point that player loses.

We implement game boards as lists of cons cells representing the
\blue{$(x,y)$} coordinates of ships. Board creation simply returns an empty
list, and adding a piece is done via \lstinline|cons|:

@@

Although \code{mark-hit} will operate on sensitive data (the game
boards), it is written without any special machinery to maintain the
secrecy of \code{board}. Protecting data {w.r.t.} policies is instead handled
automatically and implicitly by a runtime monitor.
When Alice and Bob want to play a game, they both create a
\emph{label} to protect their data. A label is unique id mapped to a policy predicate
that takes an key (e.g., the current user's name) and returns true or false to indicate
permission for the label. Alice's label is used to annotate data she wants
to be kept secret. Supposing Alice chooses to be player 1, she may use the
following label:

@@

This code illustrates label creation, policy predicates, and the first-class
nature of labels. The policy predicate $\texttt{($\lambda$ (\blue{x}) (= 1 \blue{x}))}$
grants permission to player 1 only and is associated with the dynamically generated
label l (returned and bound to \texttt{\blue{alice-label}}).
Bob would use a similar policy (but for player $2$ instead of $1$). At runtime,
the \code{let-label} form creates a label \blue{$\ell_{A}$} and binds
it to a closure for its policy predicate. When Alice wants to protect a
value, she creates a facet annotated with her label and two
\emph{branches}. The positive (left) branch represents the value as it
should appear to her, and the negative (right) to everyone else:

@@

In the above example, \blue{$\star$} (lazy failure) is used in the negative branch
to ensure execution will fail if Bob tries to observe Alice's secret gameboard.
The $\code{(obs}\ \blue{e_\ell}\ \blue{e_\text{key}}\ \blue{e_\text{fac}}\code{)}$ form takes
a label, a key, and a facted value. Explicit observation projects a single label
$\blue{e_\ell}$ in faceted value $\blue{e_\text{fac}}$ to either its positive or negative facet,
depending on whether the policy associated with $\blue{e_\ell}$ returns true for key
$\blue{e_\text{key}}$. If Bob tries to observe Alice's board, the policy predicate
will return false (from \texttt{(= 1 $\blue{2}$)}) and Alice's negative facet $\blue{\star}$ will result.

In other applications, Alice might choose a sensible default value to
reveal to others---she may even want to create a \emph{nested}
facet. For example, a social-networking application might use a
nested facet consisting of two labels for
\blue{$\ell_{\text{Friends}}$} and \blue{$\ell_{\text{Family}}$}. A user
could then present three views of her social-media profile: \bmth{p_1} to
her family, containing her phone number and other contact information,
\bmth{p_2} to her friends showing her interests, and \bmth{p_3} to
everyone else, showing only her name and email.

@@

As a game of Battleship progresses, Alice and Bob make guesses in turn, and
driver code calls the function \code{mark-hit} with each of their respective
(faceted) game boards to record the attack. However, because Alice and Bob's
game boards are both faceted values, \code{mark-hit} cannot be directly applied as
in normal execution. Instead, faceted execution ``splits'' the evaluation of
the function application over both facets, running it first on the positive
branch, then again on the negative branch. Finally, the results of
each branch are combined again to produce a new faceted value. This allows FE
to avoid needing to reason about labels and policies until an explicit observation
point where a policy is checked and a faceted value is projected to one of its
facets.

@@

Because the applied function can be stateful, faceted execution also
maintains the current privilege level in a \emph{program counter} (PC). The
program counter is a property of the current evaluation context and is used
to build facets when writes are made to the store in a privileged context.
For example, if a stateful function ``splits'' when applied on both the positive and negative
facets of a value faceted by a label $\blue{\ell}$, and on the positive branch
the function uses $\texttt{set!}$ to mutate a variable $x$ from $2$ to $3$,
FE semantics will set $x$ to $\facet{\ell}{3}{2}$ so that the value $3$
cannot be leaked from the secure context (speculative execution under the
$\blue{+\ell}$ facet). This is because, for the duration of the the app ``split'',
the evaluation context records that all values are implicitly guarded by $\blue{+\ell}$
and then $\blue{-\ell}$, respectively. If the semantics for $\texttt{set!}$ does
not make this faceting explicit, a senstive value could leak from one PC to another.
We expand upon these subtleties in Section~\ref{sec:semantics},
where we present a complete semantics for faceted execution.

@@

\begin{figure}
\begin{displaymath}
  \begin{array}{lcrcl}
    \blue{c} & \blue{\in} & \blue{const} & \blue{::=} & \blue{'() \mid \code{true} \mid \code{false} \mid \ldots} \\
    \blue{x} & \blue{\in} & \blue{var} & \blue{::=} & \langle\textit{program variables}\rangle  \\
    \blue{e} & \blue{\in} & \blue{exp} & \blue{::=} & \blue{c} \mid \blue{x} \\
    & & & \blue{\mid} & \blue{(\lambda ~ (x) ~ e) \mid (e ~ e)} \\
    & & & \blue{\mid} & \blue{(\code{box} ~ e) \mid (\code{unbox} ~ e) \mid (\code{set!} ~ e ~ e)} \\
    & & & \blue{\mid} & \blue{(\code{let-label} ~ x ~ e ~ e)} \\
    & & & \blue{\mid} & \blue{(\code{facet} ~ e ~ e ~ e)} \\
    & & & \blue{\mid} & \blue{(\code{obs} ~ e ~ e ~ e)} \\
  \end{array}
\end{displaymath}
\label{fig:syntax}
\caption{Syntax of \lang}
\end{figure}
